---
title: "Developmental changes in statistical word learning"
author: "Daniel Yurovsky & Michael C. Frank"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    theme: flatly
    code_folding: hide
    collapsed: true
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
library(knitr)
options(scipen = 1, digits = 3)
opts_chunk$set(message=FALSE, warning=FALSE, fig.align = "center", cache = TRUE)
```

```{r load_libraries, cache = FALSE, include = FALSE}
library(merTools)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(langcog)
library(directlabels)
library(lme4)
library(boot)
```

Set constants for exclusion, etc.
```{r constants}
MIN_AGE = 3
MAX_AGE = 6
MIN_LANG = 75
MIN_EXAMPLE_CORRECT = 8
RT_SDS = 3
```

## Experiment 1 ##

Read in the data and do exclusions
```{r e1_load_data}
e1_data <- read_csv("data/xsit_min.csv") %>%
  filter(age >=MIN_AGE, age < MAX_AGE) %>%
  mutate(Experiment = "Experiment 1") %>%
  mutate(trialType = factor(trialType, levels = c("Same", "Switch", "New Label")))

# Compute total number of participants
e1_total_n <- e1_data %>%
  distinct(subj) %>%
  nrow()

# Compute the number excluded due to experimenter notes
e1_excluded_n <- e1_data %>%
  distinct(subj, include) %>%
  filter(include == FALSE) %>%
  nrow()

# Drop children who do not hear English more than 75% of the time
e1_low_language_n <- e1_data %>%
  distinct(subj, include, language) %>%
  filter(include, language < MIN_LANG ) %>%
  nrow()

# compute the number excluded due to incorrect responses on example trials
e1_exclude_example_n <- e1_data %>%
  filter(include, (language > MIN_LANG | is.na(language) & setting == "Bing")) %>%
  filter(example_correct < MIN_EXAMPLE_CORRECT) %>%
  distinct(subj) %>%
  nrow()

# Compute mean and SD of log response times for exclusion 
e1_rt_exclusion_data <- e1_data %>%
  filter(include, (language > MIN_LANG | is.na(language))) %>%
  filter(example_correct >= MIN_EXAMPLE_CORRECT) %>%
  mutate(rt = log(rt), 
         mean_rt = mean(rt),
         sd_rt = sd(rt))

# Exclude trials for children who were 3 SDs above or below the mean log response times
e1_rt_exclusions <- e1_rt_exclusion_data %>%
  filter((rt < mean_rt - RT_SDS * sd_rt | rt > mean_rt + RT_SDS * sd_rt)
         & !overwrite) 

# Find the number of subjects for whom trials were excluded
e1_rt_excluded <- e1_rt_exclusions %>%
  group_by(subj) %>%
  summarise(n = n()) %>%
  group_by(n) %>%
  summarise(num_subjs = n())
  
# Get the clean dataset
e1_clean_data <-  e1_rt_exclusion_data %>%
  anti_join(e1_rt_exclusions) %>%
  mutate(age_group = floor(age*2)/2)
```

At the end of data collection, we had demographic information and data for `r e1_total_n` children. Of these, `r e1_excluded_n` were dropped due to notes from the experimenter indicating significant problems with the experiment. 

`r e1_low_language_n` children then excluded due to hearing <`r MIN_LANG`% English at home, and `r e1_exclude_example_n` additional children were dropped from further analysis due to incorrect responses on at least 1 example trial, leaving us in doubt about whether they understood the task. 

We finally performed one last round of exclusions, dropping individual trials for which the log respones time was more than 3 standard deviations away from the mean. This criterion led to excluding `r mutate(e1_rt_excluded, total = n * num_subjs) %>% summarise(total = sum(total)) %>% as.numeric` total trials, one trial for each of `r e1_rt_excluded[1,"num_subjs"]` children, and two for each of `r e1_rt_excluded[2,"num_subjs"]` children. The final sample consistend of `r nrow(e1_clean_data %>% distinct(subj))` children: 

```{r e1_data_kable}
# Print the number of participants in the final sample
e1_clean_data %>%
  distinct(subj, .keep_all = TRUE) %>%
  group_by(age_group) %>%
  summarise(n = n(), 
            prop_boys = mean(gender=="M")) %>%
  kable()
```

Let's plot some quick descriptives--how did kids do on the different trial types over age?

First, let's look at binning by year of age. 
```{r e1_descriptives}
e1_age_trial_data <- e1_clean_data %>%
  group_by(age_group, trialType, subj) %>%
  summarise(correct = mean(correct)) %>%
  group_by(age_group, trialType) %>%
  multi_boot_standard("correct", na.rm = T)

ggplot(e1_age_trial_data, aes(x = age_group, y = mean, color = trialType, 
                           label = trialType)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(.05)) + 
  geom_hline(aes(yintercept = .5), linetype = "dashed") + 
  theme_bw(base_size = 14) + 
  geom_line(aes(lty = trialType)) + 
  scale_y_continuous(limits = c(0,1),breaks=c(0,.25,.5,.75,1),
                     name = "Proportion Correct") +
  scale_x_continuous(limits=c(2.8,5.9), name = "Age Group") + 
  theme(panel.grid = element_blank()) + 
  theme(legend.position = "none", panel.grid = element_blank(),
        axis.title.x=element_text(vjust=-.5), 
        axis.title.y=element_text(vjust=1)) +
  geom_dl(method = list("last.qp", cex=1, dl.trans(x = x +.5))) + 
  scale_color_brewer(palette = "Set1")
```

Now let's try fitting a model and estimating the continuous effect
```{r e1_lm}
e1_lm <- glmer(correct ~ age * trialType + trialType * samePos + (1|subj), 
               family = "binomial", data = e1_clean_data)

kable(summary(e1_lm)$coefficients)
```

## Experiment 2 ##

Read in the data and do exclusions
```{r e2_load_data}
e2_data <- read_csv("data/xsit_me.csv") %>%
  filter(age >=3, age < 6) %>%
  mutate(Experiment = "Experiment 2") %>%
  mutate(trialType = factor(trialType, levels = c("Same", "Switch", "New Label")))

# Compute total number of participants
e2_total_n <- e2_data %>%
  distinct(subj) %>%
  nrow()

# Compute the number excluded due to experimenter notes
e2_excluded_n <- e2_data %>%
  distinct(subj, include) %>%
  filter(include == FALSE) %>%
  nrow()

# Drop children who do not hear English more than 75% of the time
e2_low_language_n <- e2_data %>%
  distinct(subj, include, language) %>%
  filter(include, language < 75 ) %>%
  nrow()

# compute the number excluded due to incorrect responses on example trials
e2_exclude_example_n <- e2_data %>%
  filter(include, (language > 75 | is.na(language))) %>%
  filter(example_correct < 8) %>%
  distinct(subj) %>%
  nrow()

# Compute mean and SD of log response times for exclusion 
e2_rt_exclusion_data <- e2_data %>%
  filter(include, (language > 75 | is.na(language))) %>%
  filter(example_correct == 8) %>%
  mutate(rt = log(rt), 
         mean_rt = mean(rt),
         sd_rt = sd(rt))

# Exclude trials for children who were 3 SDs above or below the mean log response times
e2_rt_exclusions <- e2_rt_exclusion_data %>%
  filter((rt < mean_rt - 3 * sd_rt | rt > mean_rt + 3 * sd_rt) & !overwrite) 

# Find the number of subjects for whom trials were excluded
e2_rt_excluded <- e2_rt_exclusions %>%
  group_by(subj) %>%
  summarise(n = n()) %>%
  group_by(n) %>%
  summarise(num_subjs = n())
  
# Get the clean dataset
e2_clean_data <- e2_rt_exclusion_data %>%
  anti_join(e2_rt_exclusions) %>%
  mutate(age_group = floor(age*2)/2)
```

At the end of data collection, we had demographic information and data for `r e2_total_n` children. Of these, `r e2_excluded_n` were dropped due to notes from the experimenter indicating significant problems with the experiment. 

`r e2_low_language_n` children then excluded due to hearing <`r MIN_LANG`% English at home, and `r e2_exclude_example_n` additional children were dropped from further analysis due to incorrect responses on at least 1 example trial, leaving us in doubt about whether they understood the task. 

We finally performed one last round of exclusions, dropping individual trials for which the log respones time was more than 3 standard deviations away from the mean. This criterion led to excluding `r mutate(e2_rt_excluded, total = n * num_subjs) %>% summarise(total = sum(total)) %>% as.numeric` total trials, one trial for each of `r e2_rt_excluded[1,"num_subjs"]` children, and two for each of `r e2_rt_excluded[2,"num_subjs"]` children. The final sample consistend of `r nrow(e2_clean_data %>% distinct(subj))` children:

```{r data_kable}
# Print the number of participants in the final sample
e2_clean_data %>%
  distinct(subj, .keep_all = TRUE) %>%
  group_by(age_group) %>%
  summarise(n = n(), 
            prop_boys = mean(gender=="male")) %>%
  kable()
```

```{r descriptives}
e2_age_trial_data <- e2_clean_data %>%
  group_by(age_group, trialType, subj) %>%
  summarise(correct = mean(correct)) %>%
  group_by(age_group, trialType) %>%
  multi_boot_standard("correct", na.rm = T)

ggplot(e2_age_trial_data, aes(x = age_group, y = mean, color = trialType, 
                           label = trialType)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(.05)) + 
  geom_hline(aes(yintercept = .5), linetype = "dashed") + 
  theme_bw(base_size = 14) + 
  geom_line(aes(lty = trialType)) + 
  scale_y_continuous(limits = c(0,1),breaks=c(0,.25,.5,.75,1),
                     name = "Proportion Correct") +
  scale_x_continuous(limits=c(2.8,5.9), name = "Age Group") + 
  theme(panel.grid = element_blank()) + 
  theme(legend.position = "none", panel.grid = element_blank(),
        axis.title.x=element_text(vjust=-.5), 
        axis.title.y=element_text(vjust=1)) +
  geom_dl(method = list("last.qp", cex=1, dl.trans(x = x +.5))) + 
  scale_color_brewer(palette = "Set1")
```

Now let's try fitting a model and estimating the continuous effect
```{r e2_lm}
e2_lm <- glmer(correct ~ age + trialType +  samePos + (1|subj), 
               family = "binomial", data = e2_clean_data)

kable(summary(e2_lm)$coefficients)
```

## Experiment 3 ##

Now let's look at both togehter 
```{r e1_and_2}
all_data <- bind_rows(e1_clean_data, e2_clean_data) %>%
  mutate(trialType = factor(trialType, levels = c("Same", "Switch", "New Label")))

age_trial_data <- all_data %>%
  group_by(Experiment,age_group, trialType, subj) %>%
  summarise(correct = mean(correct)) %>%
  group_by(Experiment,age_group, trialType) %>%
  multi_boot_standard("correct", na.rm = T)

ggplot(age_trial_data, aes(x = age_group, y = mean, color = trialType, 
                           label = trialType, shape = Experiment)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(.05)) + 
  geom_line(aes(lty = trialType)) + 
  geom_hline(aes(yintercept = .5), linetype = "dashed") + 
  theme_bw(base_size = 14) + 
  scale_y_continuous(limits = c(0,1),breaks=c(0,.25,.5,.75,1),
                     name = "Proportion Correct") +
  scale_x_continuous(limits=c(2.8,5.9), name = "Age Group") + 
  theme(panel.grid = element_blank()) + 
  theme(legend.position = "none", panel.grid = element_blank(),
        axis.title.x=element_text(vjust=-.5), 
        axis.title.y=element_text(vjust=1)) +
  geom_dl(method = list("last.qp", cex=1, dl.trans(x = x +.5))) + 
  scale_color_brewer(palette = "Set1")

```

Fit the big mer
```{r all_lm}
all_lm <- glmer(correct ~ age * trialType + trialType * samePos + (1|subj), 
               family = "binomial", data = all_data,
               control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e5)))

kable(summary(all_lm)$coefficients)
```